---
title: "Journal 1"
#bibliography: references.bib
author: "Ghislaine de Groot"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
#rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

<br>

------------------------------------------------------------------------

This is just an example page of your `r colorize("lab journal", "red")`. You can journal the assignments of the first day here.

------------------------------------------------------------------------

# Seminar 1

## Things I did

Replaced the picture! Enjoy my blåhaj.


# Seminar 2

## Things I did

Tried adding a second page to my website, but was unsuccessful.
Did the R Tutorial.
Read literature for the RQs.
Reformulated my research questions.
Read chapter 2.

## Literature and Research Questions 

Burris (2004) The Academic Caste System: Prestige Hierarchies in PhD Exchange Networks
- Academics gain social capital from their department (structure)
- Social capital can be combined with academic merit (e.g. research impact) (individual)
- Social capital alone has a much higher explanatory power than combined with academic merit

Hanneman (2001) The Prestige of Ph.D. Granting Departments of Sociology: A Simple Network Approach
- Prestige ranking as the behavioral realization of processes of social exchange

de Haan (2014) Sociology in the Netherlands
- Dutch Sociology departments have little differentiation

Descriptive: To what extent do Dutch sociology department differ in prestige?
Explanatory: To what extent do Dutch sociologists gain social capital and academic merit from their department?
